The \dealii{} Library, Version 9.4

### 2022-06-22 9.4/paper.tex
Adjust figure spacing.
Go through document and make some updates

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii{} release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
%  \item \texttt{AffineConstraints::make\_consistent\_in\_parallel()} allows
%  to make constraints consistent in parallel.
  \item The \texttt{DataOutResample} class interpolates values defined on one
  triangulation onto a second potentially unrelated triangulation.
  By using this class, one can output the result obtained on an
  unstructured mesh on a structured one (which might facilitate a more
  memory-efficient storage format, for example if this second
  triangulation is a uniformly refined rectangle or box), or one can create a slice in 3D.
  \item The new member function \texttt{find\_point\_owner\_rank()} of
        \texttt{parallel\allowbreak ::distributed::\allowbreak Triangulation} allows one to find the MPI
  ranks of cells containing specified points.
  It is communication-free and leverages the functionality of p4est (>v.2.2).
  Its algorithm is described in \cite{burstedde2020parallel}. This information
  enables efficient construction of the
  communication pattern used in the class \texttt{Utilties::\allowbreak MPI::\allowbreak RemotePointEvaluation}. Furthermore, this function could be used in the future to allow
  particle simulations in which particle movement is not
  limited by CFL conditions, as done in \cite{mirzadeh2016parallel}.
  \item The new function
  \texttt{GridGenerator::pipe\_junction()}
  generates a triangulation of three cone-shaped pipes that cross at a bifurcation point in any possible configuration.
  A manifold description is applied to the boundary, which can be extended into the volume via transfinite interpolation \cite{Gordon82} using the \texttt{TransfiniteInterpolationManifold} class \cite{dealII90}.
  \item A new DoF renumbering function \texttt{DoFRenumbering::support\_point\_wise()} which groups together
  shape functions by their support point. This functionality is useful in both developing nodal schemes since, e.g., the $x$, $y$, and $z$
  velocities at a point will be consecutive in the solution vector. It also improves interoperability with external libraries which expect
  data in this format.
  \item A new module \texttt{Utilities::MPI::LargeCount} which enables sending and receiving MPI messages containing more than $2^31$ objects.
  This library either uses the new MPI-4 functions, such as \texttt{MPI\_Send\_c()}, or an internal implementation of large objects for
  compatibility with MPI-3.
  \item The \texttt{FEInterfaceValues} class, which computes common quantities at the interface of two cells, has been overhauled to make it
  more consistent with the rest of the library and use more intuitive names for functions. For example,
  \texttt{FEInterfaceValues::jump\_gradient()} is now \texttt{FEInterfaceValues::jump\_in\_shape\_gradients()}. Several new
  functions, such as \texttt{FEInterfaceValues::get\_jump\_in\_function\_values()}, have also been added.
\end{itemize}
%
The changelog lists more than 100 other features and bugfixes.

@@

The matrix-free infrastructure of \dealii{} is used to solve problems
without assembling matrices, providing only the action of the matrix
instead. It has seen a number of new features in this release, the most
notable ones being:
\begin{itemize}
\item Improved support for computations with Hessians of shape functions: just like values and gradients, Hessians can be
now evaluated and integrated during matrix-free loops
both for cells and faces. This could enable, for example, writing
a matrix-free version of step-47, which solves the biharmonic equation with the discontinuous Galerkin method.
\item Cell-centric loops now also allow access to gradients and Hessians
of neighboring cells on faces. The major difficulty here was the relative orientation
of cells on unstructured meshes.
\item Users can now create their own cell batches, by providing \texttt{FEEvaluation::reinit()} a list of cell IDs. \texttt{FEEvaluation}
accesses the appropriate data and reshuffles mapping data accordingly on
the fly in order to enable vectorization over cells. The new feature is useful in many
contexts. Examples are application with sharp interfaces (e.g., two-phase flow
or shock capturing), where one needs to treat cells that are ``cut'' by
the interface in a special way. A challenge is that cell batches
might contain cut or not-cut cells, making vectorization potentially more complicated. One way to deal with such cell batches is to apply masks if
the code paths do not diverge too much
(a lot of functions of \texttt{FEEvaluation} allows this). Another way is to
categorize cells during \texttt{MatrixFree::reinit()} in such a way that mixed cell batches do not occur. However, calling  \texttt{MatrixFree::reinit()}
might be too expensive if a very dynamic system is given, which requires
recategorization in each time step. Doing this on the fly might be a cheap alternative, even if not computationally optimal during matrix-free loops.
\item Initial support for H(div)-conforming elements with Piola transform, based on Raviart--Thomas finite element spaces with the updated class \texttt{FE\_RaviartThomasNodal}. This feature is currently limited to meshes in standard orientation and affine geometries. Full support will be provided in a future release.
\item Selected matrix-free algorithms can now exploit additional data locality
  between the matrix-vector product and vector operations happening nearby in
  an algorithms. These have been added to both the
  \texttt{PreconditionChebyshev} class often used as a smoother in multigrid
  methods and the conjugate gradient implementation in \texttt{SolverCG}. To
  use these features, an operator needs to define a \texttt{vmult} operation,
  taking two additional \texttt{std::function} arguments. The first function
  defines what should be scheduled on the vector entries before the
  matrix-vector product touches them, and the second what happens
  afterwards. The new features also include a renumbering to maximize data
  locality. The theory is described in the
  contribution~\cite{kronbichler2022cg}.
\end{itemize}
Besides these new features, we improved the performance of
hanging-node-constraint evaluation on the CPU. Instead of performing
quasi-dense matrix-vector multiplications~\cite{KronbichlerKormann2012}, we now use an
approach based on in-place interpolation and sum factorization, similar
to what was already done in the GPU code~\cite{ljungkvist2017matrix}. In \cite{munch2022hn}, the algorithm is described
and performance numbers are shown, indicating a reduction
of overhead of cells with hanging nodes by a factor of ten.

Finally, we have performed a major restructuring of internals
of the \texttt{FEEvaluation} classes. This will enable us to add support for new element types, e.g., of
Raviart-Thomas and Nedelec elements.

@@

In release 9.3~\cite{dealII93}, we added support for global-coarsening multigrid in
addition to the established local-smoothing infrastructure. Global
coarsening smooths on the whole computational domain on each
multigrid level, which is obtained by coarsening the finest cells of
the next finer multigrid level.
For this purpose, we use a sequence of triangulations, and we perform
the smoothing only on their active levels. To create the sequence of
triangulations, one can use the functions \texttt{MGTransferGlobalCoarseningTools::create\_geometric\_coarsening\_sequence()}. A new version takes an
instance of \texttt{RepartitioningPolicyTools::Base} (see Subsection~\ref{sec:repartitioning}) as argument, which allows to specify the parallel
distribution of each multigrid level (in contrast to the fixed first-child policy in the case of local smoothing). Furthermore, we added support for block vectors,
fixed a number of limitations, and performed performance optimizations of
the transfer operator; particularly, the redundant copy from/to temporary vectors
has been eliminated. Now, hanging-node constraints are applied
efficiently in the same way as in the matrix-free loops (see Subsection~\ref{sec:mf}).

In \cite{munch2022gc}, the performance of the local-smoothing and global-coarsening
infrastructure of \texttt{deal.II} was compared for locally refined meshes. The results indicate that
the local definition of multigrid levels might introduce load imbalances
in the case of local smoothing so that global coarsening is favorable despite
potentially more expensive intergrid transfers. In order to judge the benefits
of one approach against the other, \texttt{deal.II} provides new functions
\texttt{workload\_imbalance()} and \texttt{vertical\_communication\_efficiency()}
in the \texttt{MGTools} namespace for the  estimation of the workload imbalance and the
vertical communication efficiency purely based on the given mesh. The publication \cite{munch2022gc} also points out that not
all types of smoothers are applicable for global coarsening due to the
presence of hanging nodes, which is a motivation to add new smoother types
to \dealii in the future.

@@

Several classes have been added to the \texttt{NonMatching} namespace to enable the use of cut finite element methods~\cite{burman_cutfem_2015}.
In the literature, these types of methods are also referred to as immersed, extended, or fictitious finite element methods.
Here, the domain, $\Omega$, is immersed in the background mesh, as illustrated in Fig.~\ref{fig:immersed-domain}.
Often, one solves for the degrees of freedom of the smallest submesh
which completely covers the domain, i.e., the blue and yellow cells of
Fig.~\ref{fig:location-to-level-set}.
The bilinear form in the weak form would then, for example, look like
\begin{equation}
  a(u,v) = (\nabla u, \nabla v)_\Omega - (\partial_n u, v)_\Gamma + \ldots
\end{equation}
Thus, when assembling on a cut call, we are
required to integrate over the part of the domain and the part of the boundary, $\Gamma = \partial \Omega$, that falls inside the cell:
$K\cap \Omega$ and $K \cap \Gamma$.
Many of the new classes that supports these operations assume that the domain is described by a level set function,
$\psi : \mathbb{R}^d \to \mathbb{R}$, such that
\begin{align}
  \Omega = \{x \in \mathbb{R}^d : \psi(x)<0\},
  \qquad
  \Gamma = \{x \in \mathbb{R}^d : \psi(x) = 0\}.
\end{align}
Specifically, the following are the key new classes and functions:
\begin{itemize}
  \item The \texttt{MeshClassifier} class identifies how the active cells and faces are located relative to the zero contour of the level set function, as illustrated in Figure~\ref{fig:location-to-level-set}. Its \texttt{location\_to\_level\_set()} function takes a cell/face and
        returns an enum, \texttt{LocationToLevelSet}, which takes values \{\texttt{inside}, \texttt{outside}, \texttt{intersected}\}.
        This information is typically needed when choosing what element a cell of the \texttt{DoFHandler} should use.

  \item The \texttt{QuadratureGenerator} class, which implements the algorithm in \cite{saye2015}, generates high-order quadrature rules for the three different regions of a \texttt{BoundingBox}, $B$, defined by the sign of the level set function:
        \begin{align}
          B \cap \Omega = \{ x\in B:  \psi(x) < 0 \},  \quad
          B \cap \Gamma = \{ x\in B:  \psi(x) = 0 \},  \quad
          \{ x\in B:  \psi(x) > 0 \}.
        \end{align}
        An example of these quadratures is shown in Figure~\ref{fig:immersed_quadratures}.
        The \texttt{FaceQuadratureGenerator} class does the same for faces.
        Furthermore, the new classes \texttt{DiscreteQuadratureGenerator} and \newline \texttt{DiscreteFaceQuadratureGenerator} can be used to generate these quadrature rules over a cell/face when the level set function lies in a finite element space: $\psi_h \in V_h$, and when the reference cell of the cell/face is a hypercube.

@@

  \item \texttt{NonMatching::FEValues} combines the functionality of
    several of the above classes to simplify assembly of linear systems. It works similarly to \texttt{hp::FEValues}:
  When calling the \texttt{reinit()} function, immersed quadrature rules are generated in the background and
  \texttt{FEValues} objects for the inside/outside region and a \texttt{FEImmersedSurfaceValues} object for the surface regions are set up internally. These can then be obtained using getter-functions and used for the assembly.
  Since the generation of immersed quadrature rules is not cheap,
  \texttt{NonMatching::FEValues} calls  \texttt{QuadratureGenerator} only if needed, i.e. if the cell is intersected. If not, already cached \texttt{FEValues} objects will be returned by the getter functions.
  Correspondingly, the class \texttt{NonMatching::FEInterfaceValues} generates
  \texttt{FEInterfaceValues} objects for assembling face terms over $F \cap \{x : \psi(x) < 0 \}$ or $F \cap \{x : \psi(x) > 0 \}$.
\end{itemize}
The new \texttt{step-85} tutorial illustrates how many of these classes work together.

@@

The Computational Geometry Algorithms Library (CGAL, \url{https://www.cgal.org/}) is a widely used
library to describe geometries and meshes \cite{cgal}. \dealii{} now
has wrappers for CGAL classes and functions, provided in the new
namespace \texttt{CGALWrappers}, and implementing functionality
spanning from mesh generation to boolean operations between \dealii{}
triangulations and cells. These wrappers are enabled only if \dealii{}
is compiled with \texttt{C++17} and, for the first time in \dealii{}'s
25-year history provides a built-in ability to generate meshes for
arbitrary geometries.

The main mesh generation function is
\texttt{GridGenerator::implicit\_function()}, which creates a \texttt{Triangulation<dim,3>} out of the zero level set of an implicit function $f$. 
For \texttt{dim==3}, the mesh consists of tetrahedra. A prototypical use case is the following, where the surface is the zero level set of Taubin's heart function $f=\bigl ( x^2 + \frac{9y^2}{4} +z^2 -1 \bigr ) -x^2 z^3 - \frac{9y^2z^3}{80}$. The resulting \texttt{Triangulation<3>} can be appreciated in Figure~\ref{fig:heart_tria}.
\begin{c++}
// Empty triangulation.
Triangulation<3, 3>  tria;
//Taubin's heart surface
ImplicitFunction                implicit_function;
CGALWrappers::AdditionalData<3> data;
data.cell_size = .05;
GridGenerator::implicit_function(tria, implicit_function, data,
                                 {0, 0, 0}, 10.0);
\end{c++}

@@

CGAL also provides us with boolean operations on meshes, as available in the
utility function
\texttt{CGALWrappers::\allowbreak{}compute\_boolean\_operation()}. The
available operations are \textit{corefinement}, \textit{intersection},
\textit{union}, and \textit{difference}. Oftentimes, boolean
operations and corefinement around the intersection produces
badly shaped elements. To overcome this issue, one can use \texttt{CGALWrappers::\allowbreak{}remesh\_surface()}. See Figure~\ref{fig:corefinement_remeshed} for a graphical example.
A possible workflow is the following:
\begin{c++}
  Triangulation<spacedim> tria0, tria1;
  GridGenerator::hyper_cube(tria0,-1.,1.);
  GridGenerator::hyper_ball(tria1,{1.,1.,1.},0.6);
  tria0.refine_global(3);
  tria1.refine_global(3);

@@

  // Compute the union of the two meshes
  CGALWrappers::compute_boolean_operation(surface_mesh0,
                            surface_mesh1,
                            BooleanOperation::compute_union,
                            out_mesh);
  // Now back to deal.II
  Triangulation<2, 3>     tria_out;
  CGALWrappers::cgal_surface_mesh_to_dealii_triangulation(out_mesh, tria_out);
  // tria_out is now a valid deal.II triangulation
  \end{c++}
  The output of the boolean operation can be seen in Fig.~\ref{fig:corefinement}, while in Fig.~\ref{fig:corefinement_remeshed}
  the same mesh has been remeshed.
  \begin{figure}
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=.2\paperwidth]{png/heart_implicit.png}
      \caption{\label{fig:heart_tria}}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[height=.14\paperheight]{png/intersection_cube_sphere_mesh.png}
      \caption{\label{fig:corefinement}}
    \end{subfigure}
    \qquad
    \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[height=.14\paperheight]{png/cube_sphere_remeshed.png}
      \caption{ \label{fig:corefinement_remeshed}}
    \end{subfigure}
    \caption{\it (a) \texttt{Triangulation<3>} of a heart surface starting from an implicit function $f$. (b) Union of the unit cube with an hyper ball. Notice how the corefinement around the intersection gives badly shaped elements. (c) The remeshed \dealii{} Triangulation.}
  \end{figure}

@@

These utilities will be the building blocks for adding functions to the \texttt{NonMatching} namespace that will assemble coupling terms like $(u,v)_{B}$, with $B$ a domain immersed in a fixed background mesh $\Omega$ and $u,v$ finite element functions on $V_h(\Omega)$, as usually happens using Nitsche's method to weakly impose boundary conditions at an interface. The same applies to coupling terms of the form $(u,q)_B$ in formulations using Lagrange multipliers, where now $q \in Q_h(B)$, with $Q_h(B)$ the space of the multiplier variable.
Notice that the most relevant difference with the quadrature rules provided by cutFEM support in \texttt{QuadratureGenerator} is that \texttt{Quadrature} objects are not created using a level-set approach, as described in the previous section, but directly with the two overlapped grids.

@@

\begin{itemize}
\item All particle data (both their identifiers and their actual data) are now stored as separate and contiguous arrays in memory, which improves prefetching of data and makes iterating over particles extremely cache efficient.
\item The choice of a list container that only includes entries for cells that contain particles means iteration is efficient, even if many cells in the domain do not contain any particles (as can be the case for discrete-element methods).
\item Creating separate arrays for each cell allows us to easily move particle IDs from one cell to another as a local operation, affecting only the two cell containers in question. We take care to reuse allocated memory to minimize the number of memory reallocations.
\item The separate cache structure that contains entries for each cell allows quick random-access to the particles of a particular cell, and also allows to quickly determine if a particular cell has particles at all.
\end{itemize}

In addition to the new storage structure we have made the following algorithmic improvements:
Determining if a particle is inside a cell after changing its position involves inverting the mapping for this cell. We have reorganized our algorithms to perform these inversions on a batch of particles in the same cell instead of particle-by-particle, which allows us to make use of vectorized instructions during the inversion.
In addition, after sorting all particles into their new cells, the arrays that store particle properties are now sorted in the same order as the particle IDs in the list of arrays, which allows for cache efficient iteration over particle properties. To avoid a costly sorting operation this operation is executed as a copy of the existing data into a new data container that replaces the existing container.

>>>>

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii{} release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
%  \item \texttt{AffineConstraints::make\_consistent\_in\_parallel()} allows
%  to make constraints consistent in parallel.
  \item The \texttt{DataOutResample} class interpolates values defined on one
  triangulation onto a second potentially unrelated triangulation.
  By using this class, one can output the result obtained on an
  unstructured mesh on a structured one (which might facilitate a more
  memory-efficient storage format, for example if this second
  triangulation is a uniformly refined rectangle or box), or one can create a slice in 3D.
  \item The new member function \texttt{find\_point\_owner\_rank()} of
        \texttt{parallel\allowbreak ::distributed::\allowbreak Triangulation} allows one to find the MPI
  ranks of cells containing specified points.
  It is communication-free and leverages the functionality of p4est (>v.2.2).
  Its algorithm is described in \cite{burstedde2020parallel}. This information
  enables efficient construction of the
  communication pattern used in the class \texttt{Utilties::\allowbreak MPI::\allowbreak RemotePointEvaluation}. Furthermore, this function could be used in the future to allow
  particle simulations in which particle movement is not
  limited by CFL conditions, as done in \cite{mirzadeh2016parallel}.
  \item The new function
  \texttt{GridGenerator::pipe\_junction()}
  generates a triangulation of three cone-shaped pipes that cross at a bifurcation point in any possible configuration.
  A manifold description is applied to the boundary, which can be extended into the volume via transfinite interpolation \cite{Gordon82} using the \texttt{TransfiniteInterpolationManifold} class \cite{dealII90}.
  \item A new DoF renumbering function \texttt{DoFRenumbering::support\_point\_wise()} which groups together
  shape functions by their support point. This functionality is useful in both developing nodal schemes since, e.g., the $x$, $y$, and $z$
  velocities at a point will be consecutive in the solution vector. It also improves interoperability with external libraries which expect
  data in this format.
  \item A new module \texttt{Utilities::MPI::LargeCount} which enables sending and receiving MPI messages containing more than $2^{31}$ objects.
  This library either uses the new MPI-4 functions, such as \texttt{MPI\_Send\_c()}, or an internal implementation of large objects for
  compatibility with MPI-3.
  \item The \texttt{FEInterfaceValues} class, which computes common quantities at the interface of two cells, has been overhauled to make it
  more consistent with the rest of the library and use more intuitive names for functions. For example,
  \texttt{FEInterfaceValues::jump\_gradient()} is now \texttt{FEInterfaceValues::jump\_in\_shape\_gradients()}. Several new
  functions, such as \texttt{FEInterfaceValues::get\_jump\_in\_function\_values()}, have also been added.
\end{itemize}
%
The changelog lists more than 100 other features and bugfixes.

@@

The matrix-free infrastructure of \dealii{} is used to solve problems
without assembling matrices, providing only the action of the matrix
instead. It has seen a number of new features in this release, the most
notable ones being:
\begin{itemize}
\item Improved support for computations with Hessians of shape functions: just like values and gradients, Hessians can be
now evaluated and integrated during matrix-free loops
both for cells and faces. This could enable, for example, writing
a matrix-free version of step-47, which solves the biharmonic equation with the discontinuous Galerkin method.
\item Cell-centric loops now also allow access to gradients and Hessians
of neighboring cells on faces. The major difficulty here was the relative orientation
of cells on unstructured meshes.
\item Users can now create their own cell batches, by providing \texttt{FEEvaluation::reinit()} a list of cell IDs. \texttt{FEEvaluation}
accesses the appropriate data and reshuffles mapping data accordingly on
the fly in order to enable vectorization over cells. The new feature is useful in several
contexts. Examples are simulations with sharp interfaces (e.g., two-phase flow
or shock capturing), where one needs to treat cells that are ``cut'' by
the interface in a special way. A challenge is that cell batches
might contain cut or not-cut cells, making vectorization of these operations potentially more complicated. Previous functionality has provided the option of masking certain cells in cell batches, which works well if
the code paths do not diverge too much. Another way is to
categorize cells during \texttt{MatrixFree::reinit()} in such a way that mixed cell batches do not occur. However, \texttt{MatrixFree::reinit()}
might be too expensive if recategorization needs to happen very frequently to follow the dynamics of a system, e.g., in each time step. Despite some overhead compared to static matrix-free loops, the new feature can be the best options in these scenarios.
\item Initial support for H(div)-conforming elements with Piola transform, based on Raviart--Thomas finite element spaces with the updated class \texttt{FE\_RaviartThomasNodal}, has been added. This feature is currently limited to meshes in standard orientation and affine geometries. Full support and performance optimizations will be provided in a future release.
\item Selected matrix-free algorithms can now exploit additional data locality
  between the matrix-vector product and vector operations happening nearby in
  an algorithm. Interfaces have been added to both the
  \texttt{PreconditionChebyshev} class (a frequently used smoother in multigrid
  methods) and the conjugate gradient implementation in \texttt{SolverCG}.
  From the user's perspective, an operator needs to define a \texttt{vmult} operation,
  taking two additional \texttt{std::function} arguments. The first function
  defines the operation to be scheduled on the vector entries before the
  matrix-vector product touches them, and the second what happens
  afterwards. The new features also include a renumbering to maximize data
  locality. The theory is described in the
  contribution~\cite{kronbichler2022cg}.
\end{itemize}
Besides these new features, we improved the performance of
hanging-node-constraint evaluation on the CPU. Instead of performing
quasi-dense matrix-vector multiplications~\cite{KronbichlerKormann2012}, we now use an
approach based on in-place interpolation and sum factorization, similar
to what was already done in the GPU code~\cite{ljungkvist2017matrix}. In \cite{munch2022hn}, the algorithm is described
and performance numbers are shown, indicating a reduction
of overhead of cells with hanging nodes by a factor of ten.

Finally, we have performed a major restructuring of internals
of the \texttt{FEEvaluation} classes. This reduces some overheads for low polynomial degrees and will enable us to add support for new element types in the future.

@@

In release 9.3~\cite{dealII93}, we added support for global-coarsening multigrid in
addition to the established local-smoothing infrastructure. Global
coarsening algorithms smoothen on the whole computational domain on each
multigrid level, which is obtained by coarsening the finest cells of
the next finer multigrid level.
For this purpose, we use a sequence of triangulations, and we perform
the smoothing only on their active levels. To create the sequence of
triangulations, one can use the functions \texttt{MGTransferGlobalCoarseningTools::create\_geometric\_coarsening\_sequence()}. A new version takes an
instance of \texttt{RepartitioningPolicyTools::Base} (see Subsection~\ref{sec:repartitioning}) as argument, which allows to specify the parallel
distribution of each multigrid level (in contrast to the fixed first-child policy in the case of local smoothing). These features have been developed and tuned for runs on supercomputer scale
with complicated coarse meshes as presented in \cite{kronbichler2021next}.
Furthermore, we added support for block vectors,
fixed a number of limitations, and performed performance optimizations of
the transfer operator; particularly, the redundant copy from/to temporary vectors
has been eliminated. Furthermore, hanging-node constraints are applied
efficiently in the same way as in the matrix-free loops (see Subsection~\ref{sec:mf}).

In \cite{munch2022gc}, the performance of the local-smoothing and global-coarsening
infrastructure of \texttt{deal.II} was compared for locally refined meshes. The results indicate that
the local definition of multigrid levels might introduce load imbalances
in the case of local smoothing so that global coarsening is favorable despite
potentially more expensive intergrid transfers. In order to judge the benefits
of one approach against the other, \texttt{deal.II} provides new functions
\texttt{workload\_imbalance()} and \texttt{vertical\_communication\_efficiency()}
in the \texttt{MGTools} namespace for the  estimation of the workload imbalance and the
vertical communication efficiency purely based on the given mesh.
% MK: I would not add this part, it does not really fit into this paper as
% there is nothing to report at this point (and we should then add references
% to the actual literature).
%The publication \cite{munch2022gc} also points out that not
%all types of smoothers are applicable for global coarsening due to the
%presence of hanging nodes, which is a motivation to add new smoother types
%to \dealii in the future.

@@

Several classes have been added to the \texttt{NonMatching} namespace to enable the use of cut finite element methods~\cite{burman_cutfem_2015}.
In the literature, these types of methods are also referred to as immersed, extended, or fictitious finite element methods.
Here, the domain, $\Omega$, is immersed in the background mesh, as illustrated in Fig.~\ref{fig:immersed-domain}.
Often, one solves for the degrees of freedom of the smallest submesh
which completely covers the domain, i.e., the blue and yellow cells of
Fig.~\ref{fig:location-to-level-set}.
The bilinear form in the weak form would then, for example, look like
\begin{equation}
  a(u,v) = (\nabla u, \nabla v)_\Omega - (\partial_n u, v)_\Gamma + \ldots
\end{equation}
Thus, when assembling on a cut cell, we are
required to integrate over the part of the domain and the part of the boundary, $\Gamma = \partial \Omega$, that falls inside the cell:
$K\cap \Omega$ and $K \cap \Gamma$.
Many of the new classes that support these operations assume that the domain is described by a level set function,
$\psi : \mathbb{R}^d \to \mathbb{R}$, such that
\begin{align}\label{eq:levelset}
  \Omega = \{x \in \mathbb{R}^d : \psi(x)<0\},
  \qquad
  \Gamma = \{x \in \mathbb{R}^d : \psi(x) = 0\}.
\end{align}
Specifically, the following are the key new classes and functions:
\begin{itemize}
  \item The \texttt{MeshClassifier} class identifies how the active cells and faces are located relative to the zero contour of the level set function, as illustrated in Figure~\ref{fig:location-to-level-set}. Its member function \texttt{location\_to\_level\_set()} takes a cell or face and
        returns an enum, \texttt{LocationToLevelSet}, with values \{\texttt{inside}, \texttt{outside}, \texttt{intersected}\}.
        This information is typically needed when choosing what element a cell of the \texttt{DoFHandler} should use.

  \item The \texttt{QuadratureGenerator} class, which implements the algorithm in \cite{saye2015}, generates high-order quadrature rules for the three different regions of a \texttt{BoundingBox}, $B$, defined by the sign of the level set function:
        \begin{align}\label{eq:boundingbox}
          B \cap \Omega = \{ x\in B:  \psi(x) < 0 \},  \quad
          B \cap \Gamma = \{ x\in B:  \psi(x) = 0 \},  \quad
          \{ x\in B:  \psi(x) > 0 \}.
        \end{align}
        An example of these quadratures is shown in Figure~\ref{fig:immersed_quadratures}.
        The \texttt{FaceQuadratureGenerator} class does the same for faces.
        Furthermore, the new classes \texttt{DiscreteQuadratureGenerator} and \texttt{DiscreteFaceQuadratureGenerator} can be used to generate these quadrature rules over a cell or face when the level set function lies in a finite element space: $\psi_h \in V_h$, and when the reference cell of the cell or face is a hypercube.

@@

  \item \texttt{NonMatching::FEValues} combines the functionality of
    several of the above classes to simplify assembly of linear systems. It works similarly to \texttt{hp::FEValues}:
  When calling the \texttt{reinit()} function, immersed quadrature rules are generated in the background and
  \texttt{FEValues} objects for the inside/outside region and a \texttt{FEImmersedSurfaceValues} object for the surface regions are set up internally. These can then be obtained using getter-functions and used for the assembly.
  Since the generation of immersed quadrature rules is not cheap,
  \texttt{NonMatching::FEValues} calls  \texttt{QuadratureGenerator} only if needed, i.e., if the cell is intersected. If not, already cached \texttt{FEValues} objects will be returned by the getter functions.
  Correspondingly, the class \texttt{NonMatching::FEInterfaceValues} generates
  \texttt{FEInterfaceValues} objects for assembling face terms over $F \cap \{x : \psi(x) < 0 \}$ or $F \cap \{x : \psi(x) > 0 \}$.
\end{itemize}
The new \texttt{step-85} tutorial illustrates how many of these classes work together.

@@

The Computational Geometry Algorithms Library (CGAL, \url{https://www.cgal.org/}) is a widely used
library to describe geometries and meshes \cite{cgal}. \dealii{} now
has wrappers for CGAL classes and functions, provided in the new
namespace \texttt{CGALWrappers}, and implementing functionality
spanning from mesh generation to boolean operations between \dealii{}
triangulations and cells. These wrappers are enabled only if \dealii{}
is compiled with \texttt{C++17} and, for the first time in \dealii{}'s
25-year history, provides a built-in ability to generate meshes for
arbitrary geometries.

The main mesh generation function is
\texttt{GridGenerator::implicit\_function()}, which creates a \texttt{Triangulation<dim,3>} out of the zero level set of an implicit function $\psi$ similar to \eqref{eq:levelset}.
For \texttt{dim==3}, the mesh consists of tetrahedra. A prototypical use case is the following, where the surface is the zero level set of Taubin's heart function $f=\bigl ( x^2 + \frac{9y^2}{4} +z^2 -1 \bigr ) -x^2 z^3 - \frac{9y^2z^3}{80}$. The resulting \texttt{Triangulation<3>} can be appreciated in Figure~\ref{fig:heart_tria}.
\begin{c++}
// Empty triangulation.
Triangulation<3, 3>  tria;
//Taubin's heart surface
ImplicitFunction                implicit_function;
CGALWrappers::AdditionalData<3> data;
data.cell_size = .05;
GridGenerator::implicit_function(tria, implicit_function, data,
                                 {0, 0, 0}, 10.0);
\end{c++}

@@

CGAL also provides us with boolean operations on meshes, as available in the
utility function
\texttt{CGALWrappers::\allowbreak{}compute\_boolean\_operation()}. The
available operations are \textit{corefinement}, \textit{intersection},
\textit{union}, and \textit{difference}. Oftentimes, boolean
operations and corefinement around the intersection produces
badly shaped mesh cells. To overcome this issue, one can use \texttt{CGALWrappers::\allowbreak{}remesh\_surface()}. Figure~\ref{fig:corefinement_remeshed} shows a graphical example.
A possible workflow is the following:
\begin{c++}
  Triangulation<spacedim> tria0, tria1;
  GridGenerator::hyper_cube(tria0,-1.,1.);
  GridGenerator::hyper_ball(tria1,{1.,1.,1.},0.6);
  tria0.refine_global(3);
  tria1.refine_global(3);

@@

  // Compute the union of the two meshes
  CGALWrappers::compute_boolean_operation(surface_mesh0,
                            surface_mesh1,
                            BooleanOperation::compute_union,
                            out_mesh);
  // Now back to deal.II
  Triangulation<2, 3>     tria_out;
  CGALWrappers::cgal_surface_mesh_to_dealii_triangulation(out_mesh, tria_out);
  // tria_out is now a valid deal.II triangulation
  \end{c++}
  The output of the boolean operation can be seen in Fig.~\ref{fig:corefinement}, while in Fig.~\ref{fig:corefinement_remeshed}
  the same mesh has been remeshed.
  \begin{figure}
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{png/heart_implicit.png}
      \caption{\label{fig:heart_tria}}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{png/intersection_cube_sphere_mesh.png}
      \caption{\label{fig:corefinement}}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.37\textwidth}
      \centering
      \includegraphics[width=\textwidth]{png/cube_sphere_remeshed.png}
      \caption{ \label{fig:corefinement_remeshed}}
    \end{subfigure}
    \caption{\it (a) \texttt{Triangulation<3>} of a heart surface starting from an implicit function $f$. (b) Union of the unit cube with an hyper ball. Notice how the corefinement around the intersection gives badly shaped elements. (c) The remeshed \dealii{} Triangulation.}
  \end{figure}

@@

These utilities will be the building blocks for adding functions to the \texttt{NonMatching} namespace that will assemble coupling terms like $(u,v)_{B}$,\marginpar{The $B$ and $\Omega$ here need to be streamlined with \eqref{eq:boundingbox}, right now they are the opposite} with $B$ a domain immersed in a fixed background mesh $\Omega$ and $u,v$ finite element functions on $V_h(\Omega)$, as usually happens using Nitsche's method to weakly impose boundary conditions at an interface. The same applies to coupling terms of the form $(u,q)_B$ in formulations using Lagrange multipliers, where now $q \in Q_h(B)$, with $Q_h(B)$ the space of the multiplier variable.
Notice that the most relevant difference with the quadrature rules provided by cutFEM support in \texttt{QuadratureGenerator} is that \texttt{Quadrature} objects are not created using a level-set approach, as described in the previous section, but directly with the two overlapped grids.

@@

\begin{itemize}
\item All particle data (both their identifiers and their actual data) are now stored as separate and contiguous arrays in memory, which improves spatial locality for better prefetching of data and makes iterating over particles extremely efficient.
\item The choice of a list container that only includes entries for cells that contain particles means iteration is efficient, even if many cells in the domain do not contain any particles (as can be the case for discrete-element methods).
\item Creating separate arrays for each cell allows us to easily move particle IDs from one cell to another as a local operation, affecting only the two cell containers in question. We take care to reuse allocated memory to minimize the number of memory reallocations.
\item The separate cache structure that contains entries for each cell allows quick random-access to the particles of a particular cell, and also allows to quickly determine if a particular cell has particles at all.
\end{itemize}

In addition to the new storage structure, we have made the following algorithmic improvements:
Determining if a particle is inside a cell after changing its position involves inverting the mapping for this cell. We have reorganized our algorithms to perform these inversions on a batch of particles in the same cell instead of particle-by-particle, which allows us to make use of vectorized instructions during the inversion using the generic scheme of~\cite{KronbichlerKormann2012}.
In addition, after sorting all particles into their new cells, the arrays that store particle properties are now sorted in the same order as the particle IDs in the list of arrays, which allows for cache efficient iteration over particle properties. To avoid a costly sorting operation this operation is executed as a copy of the existing data into a new data container that replaces the existing container.
